---
title: "Final Exam"
author: "kurella uday"
date: "2024-12-17"
output: pdf_document
---

1.1
Download the historical prices for the ticker ”SPY” from 2019-01-01 until now.
```{r}

# Load the quantmod library
library(quantmod)

# Retrieve SPY data from Yahoo Finance
getSymbols(Symbols = "SPY", from = '2019-01-01', to = '2023-05-02')

# View the first few rows of the SPY data
head(SPY)

# View the last few rows of the SPY data
tail(SPY)

```

1.2
Calculate the daily log returns for SPY using the adjusted close prices.
```{r}
# Convert the SPY data to a data frame
SPY <- data.frame(SPY)

# Extract the adjusted closing prices
SPY.price <- SPY$SPY.Adjusted

# Calculate the log of adjusted prices
SPY.log.price <- log(SPY.price)

# Calculate the log returns as the difference of log prices
SPY.log.return <- diff(SPY.log.price)

# Display the first few rows of the log returns
head(SPY.log.return)

```

1.3
Plot the daily log returns in red line.

```{r}
# Plot the log returns with a line type and red color
plot(SPY.log.return, 
     type = "l",               # Line plot
     col = "red",              # Line color
     main = "Log Returns of SPY", # Title of the plot
     xlab = "Time",            # X-axis label
     ylab = "Log Return",      # Y-axis label
     lwd = 2,                  # Line width
     col.main = "black",       # Title color
     col.lab = "blue",         # Axis label color
     cex.lab = 1.2)            # Axis label size

```

2.1
Calculate the skewness and kurtosis of the SPY daily log return from Question 1, for both adjusted and unadjusted ones. (See page 21 and 23 of L6 and the corresponding HW problems)

```{r}
# Function for calculating skewness
skewness <- function(x, adjusted = FALSE){
  # Calculate the mean once for efficiency
  x_mean <- mean(x)
  
  # Calculate third central moment (m3) and second central moment (m2)
  m3 <- mean((x - x_mean)^3)
  m2 <- mean((x - x_mean)^2)
  
  # Calculate unadjusted skewness
  m_tilde_3 <- m3 / m2^(3/2)
  
  # Adjusted skewness if requested
  if(adjusted == FALSE){
    return(m_tilde_3)
  } else {
    n <- length(x)
    m_tilde_3_adjusted <- (sqrt(n * (n - 1)) / (n - 2)) * m_tilde_3
    return(m_tilde_3_adjusted)
  }
}

# Function for calculating kurtosis
kurtosis <- function(x, adjusted = FALSE){
  # Calculate the mean once for efficiency
  x_mean <- mean(x)
  
  # Calculate fourth central moment (m4) and second central moment (m2)
  m2 <- mean((x - x_mean)^2)
  m4 <- mean((x - x_mean)^4)
  
  # Calculate unadjusted kurtosis
  m_tilde_4 <- m4 / m2^2
  
  # Adjusted kurtosis if requested
  if(adjusted == FALSE){
    return(m_tilde_4)
  } else {
    n <- length(x)
    m_tilde_4_adjusted <- ((n - 1) / ((n - 2) * (n - 3))) * ((n + 1) * m_tilde_4 - 3 * (n - 1)) + 3
    return(m_tilde_4_adjusted)
  }
}

# Test the skewness and kurtosis functions on SPY log returns
skew_result_unadjusted <- skewness(SPY.log.return, adjusted = FALSE)
skew_result_adjusted <- skewness(SPY.log.return, adjusted = TRUE)

kurt_result_unadjusted <- kurtosis(SPY.log.return, adjusted = FALSE)
kurt_result_adjusted <- kurtosis(SPY.log.return, adjusted = TRUE)

# Print results
cat("Unadjusted Skewness: ", skew_result_unadjusted, "\n")
cat("Adjusted Skewness: ", skew_result_adjusted, "\n")
cat("Unadjusted Kurtosis: ", kurt_result_unadjusted, "\n")
cat("Adjusted Kurtosis: ", kurt_result_adjusted, "\n")

```

2.2
Report the results in 2.1 using a 2×2 table (either data frame or matrix) such that: The column names are ”SPY.skewness” and ”SPY.kurtosis”. And the row names are ”Unadjusted” and ”Adjusted”.

```{r}

# Create a data frame with adjusted and unadjusted skewness and kurtosis
x <- data.frame(
  SPY.skewness = c(skewness(SPY.log.return, adjusted = TRUE), skewness(SPY.log.return, adjusted = FALSE)),
  SPY.kurtosis = c(kurtosis(SPY.log.return, adjusted = TRUE), kurtosis(SPY.log.return, adjusted = FALSE))
)


rownames(x) <- c("Adjusted", "Unadjusted")

# Print the resulting data frame
print(x)

```

3.1
Download options prices for ticker ”SPY” for all expiration dates.

```{r}
library(quantmod)
SPY.option<- getOptionChain("SPY")
head(SPY.option)



```

3.2
For calls and puts of each expiration date, add a column of ”Price”, which
is the average of “Bid” and “Ask”.

```{r}
SPY.option<- getOptionChain("SPY", NULL)
for (i in 1:length(SPY.option)) {
  
  SPY.option[[i]]$calls$Price<-0.5*(SPY.option[[i]]$calls$Bid + SPY.option[[i]]$calls$Ask)
  SPY.option[[i]]$puts$Price<-0.5*(SPY.option[[i]]$puts$Bid + SPY.option[[i]]$puts$Ask)
}
#head(SPY.options)
```

3.3
Choose 3 expiration date for put options, plot volatility smiles (Strike in x-axis and ImpliedVol in y-axis, similar to call smiles on page 22 of L9).

```{r}
for (i in 1:length(SPY.option)) {
SPY.option[[i]]$calls$Price <- 0.5*(SPY.option[[i]]$calls$Bid +SPY.option[[i]]$calls$Ask)
SPY.option[[i]]$puts$Price <- 0.5*(SPY.option[[i]]$puts$Bid + SPY.option[[i]]$puts$Ask)
}
sigma <- 0.2
bisection.new <- function(f, a, b, tol = 0.001, N.max = 100){
f.a <- f(a)
f.b <- f(b)
if(f.a*f.b > 0){
}else if(f.a == 0){
return(a)
}else if(f.b == 0){
return(b)
}else if(is.na(f.a*f.b)){
return(NA)
}
for(n in 1:N.max){
c <- (a+b)/2
f.c <- f(c)
if(f.c == 0 || abs(b - a) < tol){
break
}
if(f.a*f.c < 0){
b <- c
f.b <- f.c
}else{
a <- c
f.a <- f.c
}
}
return(c)
}
bs.call <- function(S0, K, T1, sigma, r){
d1 <- (log(S0/K) + (r+0.5*sigma * sigma)*T1)/(sigma*sqrt(T1))
d2 <- d1 - sigma*sqrt(T1)
S0
}

bs.put <- function(S0, K, T1, sigma, r){
d1 <- (log(S0/K) + (r+0.5*sigma * sigma)*T1)/(sigma*sqrt(T1))
d2 <- d1 - sigma*sqrt(T1)
-S0*pnorm(-d1) + exp(-r*T1)*K*pnorm(-d2)
}
Vega <- function(S0, K, T1, sigma, r){
d1 <- (log(S0/K) + (r+0.5*sigma * sigma)*T1)/(sigma*sqrt(T1))
sqrt(T1)*S0*dnorm(d1)
}
dprice.diff <- function(sigma)Vega(S0,K,T1,sigma,r)
implied.vol<- function(S0, K, T1, r, price,type, method="bisection"){
price.diff.call <- function(sigma)bs.call(S0, K, T1, sigma, r) - price
price.diff.put <- function(sigma)price - bs.put(S0, K, T1, sigma, r)
if(method == "bisection"&&type =="call"){
return(bisection.new(price.diff.call, 0.01, 5))
}else{
return(bisection.new(price.diff.put, 0.01, 5))
}
if(method == "Newton-Raphson"&&type=="call"){
dprice.diff <- function(sigma)Vega(S0, K, T1, sigma, r)
return(Newton_Raphson(price.diff.call, dprice.diff, 0.25))
}else{
return(Newton_Raphson(price.diff.put, dprice.diff, 0.25))
}
}
```

3.4
Choose 3 expiration date for put options, plot volatility smiles (Strike in x-axis and ImpliedVol in y-axis, similar to call smiles on page 22 of L9).

```{r}
SPY.expiration <- names(SPY.option)
T.vec <- (as.Date(SPY.expiration, "%b.%d.%Y") - Sys.Date()) / 365
T.vec <- as.numeric(T.vec)
SPY.S0 <- getQuote("SPY")$Last
r <- 0.07 * 0.01

for(i in 1:length(SPY.option)){
  # Compute the average price for calls
  SPY.option[[i]]$calls$Price <- 0.5 * (SPY.option[[i]]$calls$Bid + SPY.option[[i]]$calls$Ask)
  
  for(j in 1:nrow(SPY.option[[i]]$calls)){
    # Ensure that no NA values are passed to the implied volatility calculation
    if(!is.na(SPY.option[[i]]$calls$Price[j]) && !is.na(SPY.option[[i]]$calls$Strike[j])){
      SPY.option[[i]]$calls$impliedVol[j] <- implied.vol(SPY.S0, SPY.option[[i]]$calls$Strike[j], T.vec[i], r, SPY.option[[i]]$calls$Price[j], "calls")
    } else {
      SPY.option[[i]]$calls$impliedVol[j] <- NA  # Assign NA if the price or strike is missing
    }
  }
  
  SPY.option[[i]]$calls <- SPY.option[[i]]$calls[c("Bid", "Ask", "Strike", "Price", "impliedVol")]
}

for(i in 1:length(SPY.option)){
  # Compute the average price for puts
  SPY.option[[i]]$puts$Price <- 0.5 * (SPY.option[[i]]$puts$Bid + SPY.option[[i]]$puts$Ask)
  
  for(j in 1:nrow(SPY.option[[i]]$puts)){
    
    if(!is.na(SPY.option[[i]]$puts$Price[j]) && !is.na(SPY.option[[i]]$puts$Strike[j])){
      SPY.option[[i]]$puts$impliedVol[j] <- implied.vol(SPY.S0, SPY.option[[i]]$puts$Strike[j], T.vec[i], r, SPY.option[[i]]$puts$Price[j], "puts")
    } else {
      SPY.option[[i]]$puts$impliedVol[j] <- NA  # Assign NA if the price or strike is missing
    }
  }
  
  SPY.option[[i]]$puts <- SPY.option[[i]]$puts[c("Bid", "Ask", "Strike", "Price", "impliedVol")]
}

# Plot implied volatilities for puts
plot(NA, xlim = c(300, 500), ylim = c(0, 0.4), xlab = "Strike", ylab = "Implied Volatility")
lines(SPY.option[[1]]$puts$Strike, SPY.option[[1]]$puts$impliedVol, col = "red")
lines(SPY.option[[2]]$puts$Strike, SPY.option[[2]]$puts$impliedVol, col = "green")
lines(SPY.option[[3]]$puts$Strike, SPY.option[[3]]$puts$impliedVol, col = "blue")
legend("bottomleft", SPY.expiration[c(6, 9, 17)], fill = c("red", "green", "blue"))

```

3.5
Keep fields “Strike”,“Bid”,“Ask”, “Price”, and “ImpliedVol” and save the calls and puts of each expiration date in .csv file. Submit one of the .csv file also.
```{r}

print(getwd())

setwd("C:\\Users\\K.uday\\Desktop")
Sys.Date()

(today <- format(Sys.Date(), "%Y_%m_%d"))
(Exp <- names(SPY.option))

Exp <- as.Date(Exp, format = "%b.%d.%Y")
for (i in 1:length(Exp)) {
write.csv(SPY.option[[i]]$calls,
file = paste0("data",today,"Exp",Exp[i],"calls.csv"))
write.csv(SPY.option[[i]]$puts,
file = paste0("data",today,"Exp",Exp[i],"puts.csv"))
}

```